Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-01 manuscript.Rmd
Reorganize introduction
Edits to Social Media and Social Movements section

Prior to the advent of social media, people coordinated their protest behavior
using phone tree calling sheets and mimeographed flyers,
through in-person meetings of social organizations and churches, and through
word of mouth. They gathered together and hand-made protest signs that reflected
their agendas. As McAdam found, people were more likely to join movements when
they had a personal connection [@mcadam1986recruitment; @mcadam1993specifying]: 
Historically, strong ties have been
strong predictors of involvement in high-risk protest activity. 
However, the rise of the Internet and its decentralized online networks
has upended theories about recruitment to---and participation in---social
movements. Strong interpersonal ties have yielded to weak impersonal ties that
strongly predict involvement in social movements [@steinert2017spontaneous]. Research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to ``big data'' analyses of language on social media. 
*(Probably need to wrap up with a sentence here.)*

@@

In the Internet age, social media replicates and amplifies
this phenomenon during moments of social unrest. Amid the cacophony of messages
during political upheaval---such as the Arab Spring---consistent themes emerge
through hashtags, retweets, and the repeated mentions of concepts such as
regime, security, demonstration, and human rights. The cohesiveness of these
themes (i.e., whether movement participants are metaphorically singing in one voice) 
reflects the real-world events happening contemporaneously with collective action in
social movements or civil conflicts.

@@

## Social Media and Social Movements
During the Arab Spring, social media provided real-time information about the
events transpiring in countries across the Middle East and North Africa. Our
understanding about the role of social media in social movements is still
evolving. Some argue that these mass movements would not have been possible
without efficient large-scale coordination, both physically on the ground and
virtually on social media platforms. Along with Facebook, Twitter has been
credited as essential to the Arab Spring---both for protesters (as a means to
coordinate with one another and to communicate with the world) and for the
government response (to monitor and track protestors to crack down on dissent)
[@comninos2011twitter]. The events of Arab Spring have been thoroughly discussed
within the political science literature [e.g., @lutterbeck2013arab;
@koehler2017political], but to our knowledge, the direct *coupling* between
Syrian Twitter cohesion and international action coordination has not been
investigated. Others provide skepticism about the effect of social media on
collective action [@Seay_2014].

In Foreign Affairs, Anderson writes that, “The important story about the 2011
Arab revolts in Tunisia, Egypt, and Libya is not how the globalization of the
norms of civic engagement shaped the protesters' aspirations. Nor is it about
how activists used technology to share ideas and tactics. Instead, the critical
issue is how and why these ambitions and techniques resonated in their various
local contexts [@Anderson_2011].” Malcolm Gladwell, in highlighting the notions
of strong and weak ties, @Gladwell_2010 and
@McAdam_1986, similarly expressed hesitation about the role of online
organizing, arguing that real-life strong ties outweigh online weak ties and
that participation in online activism only requires scant sacrifices whereas
recruitment to higher-risk activities is done through in-person contact, namely
strong ties [@Gladwell_2010; [@McAdam_1986]. Interestingly, these views may
actually help to make the case for bridging the gap between social media and
social mobilization, by suggesting the strength of social ties in an online
environment is inextricably linked to the context and ambitions of the actors.
Granovetter notes that while strong ties tend to bind smaller groups, weak ties
tend to bind groups across distances [@Granovetter_1973].

Social media generates critical information about events, such as whether or not
a demonstration will take place and what transpires when it does, how many
people participate, and how the state, the military, and the security apparatus
respond. Similarly, these media platforms alert us to more global shifts within
a conflict, such as when fighting moves to a new city or region, the number of
casualties and wounded, and the types of weaponry used by parties to the
conflict [@Zeitzoff_2017]. Social media is not entirely event-driven, however.
In monitoring social media channels, government forces may respond to the
perception that citizens are gaining strength and sympathy. Governments have
been shown to be quite tolerant of dissent so long as it does not metamorphosize
into collective action [@King_Pan_Roberts_2013]. Additionally, the insulated
nature of online social platforms allows valuable social dynamics to be detailed
at unprecedented resolution and scale. As such, the digital traces left by
individuals and collectively as groups are regarded as a valuable source of
information that can offer powerful insights into states of social
disequilibrium, up-risings, and escalations into civil wars.

Citizens communicate with each other with increased transparency due to the
ubiquitous availability of social media such as low-cost and egalitarian
platforms like Twitter. Citizens are able to access these tools via mobile
phones and computers with Internet connectivity, providing a stream of real-time
shared information. This facilitates studying the degree of cohesion between
speakers, and the alignment between that cohesion and their collective political
behavior. Citizens can tweet the location of events like protests and military
incursions, the degree of state repression, and the number of casualties. In
kind, the government, interested in remaining in power, monitors social media
platforms to ascertain emerging threats, such as where new groups are forming
and where existing groups are strengthening, where gatherings are held, and who
the idea entrepreneurs are. An example of a leader manifesting his interest in
such public conversation is given in a speech of June 3, 2012 by Bashar
al-Assad:

“I know that I have been away from the media for a long time, and I have missed
having direct contact with the citizens, but I have always been following up
with the daily occurrences and gathering the information so that my speech can
be built on what is said by the street.”

With few exceptions, Assad chose to keep social media platforms up and running.
Gohdes (2015) notes that regimes can attempt to thwart the growth of social
protest online and in the streets by shutting down the Internet [@Gohdes_2015].
During these times, however, she finds that military repression increases.
However, government information about the citizens decreases, and as such, many
regimes have elected to maintain Internet connectivity as a means of monitoring
and surveilling their citizens.

>>>>

Prior to the advent of social media, people coordinated their protest behavior
using phone tree calling sheets and mimeographed flyers,
through in-person meetings of social organizations and churches, and through
word of mouth. They gathered together and hand-made protest signs that reflected
their agendas. As McAdam found, people were more likely to join movements when
they had a personal connection [@mcadam1986recruitment; @mcadam1993specifying]: 
Historically, strong ties have been
strong predictors of involvement in high-risk protest activity.

@@

However, the rise of the Internet and its decentralized online networks
has upended theories about recruitment to---and participation in---social
movements. Strong interpersonal ties have yielded to weak impersonal ties that
strongly predict involvement in social movements [@steinert2017spontaneous]. Research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to ``big data'' analyses of language on social media.

Today, social media replicates and amplifies
this phenomenon during moments of social unrest. Amid the cacophony of messages
during political upheaval---such as the Arab Spring---consistent themes emerge
through hashtags, retweets, and the repeated mentions of concepts such as
regime, security, demonstration, and human rights. The cohesiveness of these
themes (i.e., whether movement participants are metaphorically singing in one voice) 
reflects the real-world events happening contemporaneously with collective action in
social movements or civil conflicts.

@@

## Social Media and Social Movements
During the Arab Spring, social media provided real-time information about the
events transpiring in countries across the Middle East and North Africa. Our
understanding about the role of social media in social movements is still
evolving. Some argue that these mass movements would not have been possible
without efficient large-scale coordination, both physically on the ground and
virtually on social media platforms.

Along with Facebook, Twitter has been
credited as essential to the Arab Spring---both for protesters (as a means to
coordinate with one another and to communicate with the world) and for the
government response (to monitor and track protestors to crack down on dissent)
[@comninos2011twitter]. The events of Arab Spring have been thoroughly discussed
within the political science literature [e.g., @lutterbeck2013arab;
@koehler2017political], but to our knowledge, the direct *coupling* between
Syrian Twitter cohesion and international action coordination has not been
investigated. In fact, some have been skeptical about the effect of social media on
collective action at all [@seay2014slacktivism].

This skepticism can be summarized in a quote by Lisa Anderson:

> The important story about the 2011
Arab revolts in Tunisia, Egypt, and Libya is not how the globalization of the
norms of civic engagement shaped the protesters' aspirations. Nor is it about
how activists used technology to share ideas and tactics. Instead, the critical
issue is how and why these ambitions and techniques resonated in their various
local contexts. [anderson2011demystifying]

Others have similarly expressed hesitation about the role of online
organizing. Some have argued that real-life strong ties outweigh online weak ties and
that the power of the strong ties through in-person contact is critical
to recruiting people for higher-risk activities 
[@gladwell2010revolution; @mcadam1986recruitment].

Interestingly, these views may
actually help to make the case for bridging the gap between social media and
social mobilization. Such views suggest the strength of social ties in an online
environment is inextricably linked to the *context* and *ambitions* of the actors.
Moreover, research suggests that weak ties still retain considerable political
power: While strong ties tend to bind smaller groups, weak ties
tend to bind groups across distances [@granovetter1977strength].

Social media generates critical information about events, including whether and how
a demonstration will take place, how many
people participate, and how those in power (e.g., state, military, security apparatus)
respond. Similarly, these media platforms alert us to more global shifts within
a conflict, such as when fighting moves to a new city or region, the number of
casualties and wounded, and the types of weaponry used by parties to the
conflict [@zeitzoff2017social]. Social media is not entirely event-driven, however.
In monitoring social media channels, government forces may respond to the
perception that citizens are gaining strength and sympathy. Governments have
been shown to be quite tolerant of dissent so long as it does not metamorphose
into collective action [@king2013censorship].

Citizens communicate with each other with increased transparency due to the
ubiquitous availability of social media such as low-cost and egalitarian
platforms like Twitter. Citizens are able to access these tools via mobile
phones and computers with Internet connectivity, providing a stream of real-time
shared information. Citizens can share the location of events like protests and military
incursions, the degree of state repression, and the number of casualties. In
turn, the government---interested in remaining in power---monitors social media
platforms to ascertain emerging threats [@comninos2011twitter], 
such as where new groups are forming
and where existing groups are strengthening, where gatherings are held, and who
the idea entrepreneurs are. An example of a leader 
expressing his interest in
such public conversation is given by Bashar al-Assad in a speech on June 3, 2012:

>I know that I have been away from the media for a long time, and I have missed
having direct contact with the citizens, but I have always been following up
with the daily occurrences and gathering the information so that my speech can
be built on what is said by the street.

With few exceptions, Assad chose to keep social media platforms up and running.
Gohdes [-@gohdes2015pulling] notes that regimes can attempt to thwart the growth of social
protest online and in the streets by shutting down the Internet.
During these times, however, she finds that military repression increases.
However, government information about the citizens decreases; as such, many
regimes have elected to maintain Internet connectivity as a means of monitoring
and surveilling their citizens.

At the same time, the insulation
of online social platforms allows researchers to detail real-world social dynamics
at unprecedented resolution and scale [@paxton2017finding]. The digital traces left by
individuals and collectively as groups are regarded as a valuable source of
information that can offer powerful insights into states of social
disequilibrium, uprisings, and escalations into civil wars.
By examining these conversations, researchers can study the degree of cohesion between
speakers and the alignment between that cohesion and their collective political
behavior.
